%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%   SETHEO MANUAL
%	(c) J. Schumann, O. Ibens
%	TU Muenchen 1995
%
%	%W% %G%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The First Order Predicate Logic Syntax}\label{sec:pl1-syntax}

In\footnote{This text is from the OLD SETHEO MANUAL.}
 this part we describe the syntax of the input language, which is
accepted by the formula translator {\bf plop}.
A formula in that notation has to be in a file with the extension {\bf .pl1}.

Because the grammar of first order logic is well-known it seems
useless to repeat an inductive definition of the concept 
{\em ``formula''}. We rather represent the special PLOP
requirements i.e. the differences to the usual
standard in giving some comments on the individual
elements of the accepted language

\paragraph{Variables}
Variables are strings starting with a capital letter followed 
by arbitrary letters, digits, or underscore.

\paragraph{Constants}
Predicate-, function-, and object parameters are strings starting 
with a lower case letter followed by lower case letters, digits, or underscore. 
Predicate- and function arguments are to be included in round brackets.
Multiple arguments are to be separated by a comma ``,''.

\paragraph{Propositional Connectives}
The symbols for the propositional connectives are:

\begin{center}
\begin{tabular}{llll}
(1) &  {\tt $\sim$}  &   - &  not, \\
(2)  &	 {\tt \& } & -  & and, \\
(3)   &{\tt ; } & -  & or, \\
(4)   &{\tt -> } & -  & if ... then, \\
(5) & 	 {\tt <->}  & -  & iff. 
\end{tabular}                  
\end{center}

The binding preference is given by this order:
1 $>$ 2 $>$ 3 $>$ 4 $>$ 5 . 
Equal connectives are left associative i.e.,
\tt 
\mbox{a -> b -> c} \rm is to be read as \tt \mbox{((a -> b) -> c)} 
\rm .

\paragraph{Quantifiers}
The quantifiers are represented by
``{\tt
forall}'' and ``{\tt exists}''.
These symbols must not occur as parts of other names like constants
or variables.
The scope of a quantifier extends over the following propositionally 
complete subformula and ends there.
E.g. at the formula
\[\tt 
\mbox{forall X a -> b(X)}  \\
\rm
\]
the scope of the quantifier does not include {\tt b(X)}.
To bind {\tt X} by this quantifier one has to write as usual with parenthesis
\[\tt 
\mbox{forall X (a -> b(X)).}\\
\rm
\]
Occurences of free variables are understood by PLOP as bound
from the outside (universal closure).

\paragraph{Brackets}
Parenthesis are expressed by the usual {\em round brackets} only.
{\em Square brackets} must not be used in the whole input formula .

\paragraph{Special Characters}
All characters with ASCII value $<$ 32 are ignored and will be deleted
during the transformation.
The {\em space-character} can be omitted if the meaning is not affected and if
there cannot result any ambiguities. 
E.g. 
\[\tt 
\mbox{forallX} \\
\rm
\]
is recognized as a quantifier and a variable. 
In other cases a space character must be set, e.g.
\[\tt 
\mbox{forall X a(X) }\\
\rm
\]
cannot be written as 
\[ \tt 
\mbox{forall Xa(X).}\\
\rm
\]


Other characters as mathematical function symbols (especially in
infix notation) are not recognized.

\paragraph{Comments}
Comments are enclosed in C--like convention in ``{\tt /*}'' and ``{\tt */}''
and can extend over several lines. Nested comments are not allowed.
All comments are removed during the conversion.
