%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%   SETHEO MANUAL
%	(c) J. Schumann, O. Ibens
%	TU Muenchen 1995
%
%	%W% %G%
%
% BUILT-IN:	
% SCCS:		%W% %G%
% AUTHOR:	J. Schumann
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\$set{\em bound\/}/1}

This section describes a number of built-ins which allow to destructively 
set the current (or maximal) values of the search bounds.
All assignements to the bounds are backtrackable, i.e., their effect
is undone as soon a a backtracking occurrs over the corresponding built-in.
\begin{description}
\item[Synopsis:]
	{\tt \$set{\em bound}(N)}
\item[Parameters:]\ \\[-0.5cm]
	\begin{description}
	\item[{\tt N}] term which must evaluate to a non-negative number.
	\end{description}
\item[Low Level Name:]
	{\tt set{\em bound}}
\item[Result:]\ \\
\end{description}

\vspace*{0.5cm}
\noindent
{\bf Description.}
A built-in of this group sets the current or maximal value
of a search-bound.
\begin{description}
\item[{\tt \$setdepth(N)}]
sets the bound to the maximum tableau depth
            (allowed for a final tableau) minus current depth.
\item[{\tt \$setinf(N)}]
   sets the number
            of inferences.
\item[{\tt \$setlocinf(N)}]
   sets the number
            of local inferences.
\item[{\tt \$setmaxinf(N)}]
Obtains the maximum number
            of inferences (allowed for a final tableau).
\end{description}

\vspace*{0.5cm}
\noindent
{\bf Side-effects.}
These built-in effect global registers of the SETHEO Abstract Machine
\SAM.
Therefore, a wrong usage of these built-ins can have undesired effects
and can destroy the completeness of the proof procedure.

\vspace*{0.5cm}
\noindent
{\bf Notes.}
These built-ins are used to adapt the iterative deepening, or to implement
Meta-interpreters. For details and examples refer to Section~\ref{chap:How-to}.

\vspace*{0.5cm}
\noindent
{\bf Example.}
\begin{verbatim}
p(X) <- getdepth(D),D1 is D+1,setdepth(D1),q(X),setdepth(D).
\end{verbatim}

When this clause is called, then its subgoal {\tt q} can use  the same
resources (here, the depth), as the clause itsself.
