\subsection{Filename Conventions}

All files used within the SETHEO system carry specific file name
extensions. These extensions are mandatory and are {\em not} given
when a SETHEO command (see Section~\ref{chap:2},\ref{chap:3})
is issued.
In the following we list all possible file name extensions and describe
their contents. Formal definitions of the syntax is given in 
Chapter~\ref{ch:syntax}.

\begin{description}
\item[{\em file}.pl1]
Files with this extension contain formulae in first order predicate
logic in mathematical syntax. Formulae can contain the standard
operators (like $\vee, \wedge,\rightarrow,\leftrightarrow$) and
existential and universal quantifiers. An ASCII representation
of the quantors is given, e.g., {\tt forall} for $\forall$.
A detailed description and formal definition is given in
Section~\ref{sec:pl1-syntax}.
In the current version of SETHEO, this notation is not supported.
Note, however, that this syntax is very similar to the input syntax
of OTTER and can be converted easily (see Section~\ref{sec:plop2otter}).

\item[{\em file}.lop]
Files with this extension contain formulae and/or logic programs
written in LOP syntax. This syntax comprises the default for the
SETHEO system.
A detailed description and the syntax definition is given in
Section~\ref{sec:lop-syntax}. Built-in predicates which can be used
for logic programming are listed in Chapter~\ref{chap:builtins}.
Files with the extension {\bf .lop} comprise the input of the compiler
{\bf inwasm}.

\item[{\em file}\_pp.lop]
Files with this extension are produced by the compiler {\bf inwasm}
when it is invoked with the command-line option {\tt -lop}.
This file contains the formula after preprocessing. Clauses are fanned
into contra-positives, constraints are added to the clauses
(if selected), and the weak-unification graph is given.
Such a file can be read again by the compiler {\bf inwasm}.
Thus, manual (or automatic) modifications of the formula between
preprocessing and the proof attempt can be accomplished.

\item[{\em file}.out.lop]
Files with this extension are produced by the {\sc Delta} iterator
when given {\em file\/} as input.
This file contains (in LOP syntax) the original formula (as found
in {\em file\/}{\bf .lop}) with the newly generated unit clauses
appended to it.

\item[{\em file}.s]
Files with this extension contain \SAM\ assembler statements and
are the input of the assembler {\bf wasm}.
Assembler files are generally generated by the compiler {\bf inwasm}.
Nevertheless, assembler files are (to some extend) human-readable
and can be modified (for prototypical implementation issues).

\item[{\em file}.hex]
Such files are generated by the assembler {\bf wasm} and contain the
binary (i.e., hexadecimal) representation of \SAM-instructions and
\SAM's symbol table. {\bf .hex} files are directly loaded by the
SETHEO Abstract Machine \SAM (command {\tt sam}.

\item[{\em file}.tree]
The prover {\tt sam} generates files with that extension,
if a proof has been found or upon request. Such files contain
one or more Model Elimination Tableaux which then can be displayed
graphically by the proof tree viewer {\bf xptree}.
The syntax of each tableau corresponds to a PROLOG term. Therefore,
tableaux can be read in by a PROLOG system or (after adding
predicate symbols to it) by the compiler {\bf inwasm}.
The syntax of the tree-files are described in Section~\ref{sec:tree-syntax}.
\item[{\em file}.log]
This file is generated by the prover {\tt sam}. It contains useful
information about the current run and comprises
a copy of the data displayed on stdout.
Besides command-line switches and a log of the iterative deepening, this
file contains statistical information which is printed when the
\SAM\ stops (successfully or not).
\end{description}


Two further types of files are used within the SETHEO-system which
do not carry specific extensions.
The \SAM-output file is opened by the built-in predicate {\tt tell/1}
and contains all subsequent output as produced by other built-ins
(e.g., {\tt write} and {\tt dumplemma}).

The {\bf xptree} operator file contains  a translation table
of all operators together with their binding power. Its syntax is
defined in Section~\ref{sec:xtab-syntax}.
